19Q1

### 2019-03-27 19Q1.Rmd

## Denise's GCN dataframes

### AUC

### Summary figures

@@

```{r}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
ccle_poz <- merge(ccl_histologies, ccle_poz, by = "Cell_line", all.y = TRUE)
ccle_poz$InCCLE <- "1"
ccle_poz$InCTRP <- ""
ccle_poz$InGDSC <- ""

@@

auc_unmatched <- bind_rows(ccle_poz, ctrp_poz, gdsc_poz)
# auc_unmatched$AUC_EC50 <- as.numeric(as.character(auc_unmatched$AUC_EC50))
auc_unmatched$AUC_IC50 <- as.numeric(as.character(auc_unmatched$AUC_IC50))
# auc_unmatched$sample_ID <- with(auc_unmatched, paste(Database, Cell_line, Histology, sep = "_"))
auc_unmatched$Drug_name_Database <- with(auc_unmatched, paste(Drug_name, Database, sep = "_"))

@@

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_unmatched_wide <- dcast(auc_unmatched, Cell_line + Histology + CCLE_ccl_name + Broad_ID + InCCLE + InCTRP + InGDSC ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_unmatched_wide[auc_unmatched_wide == 1000000] <- ""
auc_unmatched_wide <- auc_unmatched_wide %>% group_by(Cell_line, Histology, CCLE_ccl_name, Broad_ID) %>% summarise_all(funs(trimws(paste(., collapse = ''))))
auc_unmatched_wide$InCCLE[auc_unmatched_wide == ""] <- 0
auc_unmatched_wide$InCTRP[auc_unmatched_wide == ""] <- 0
auc_unmatched_wide$InGDSC[auc_unmatched_wide == ""] <- 0
auc_unmatched_wide[auc_unmatched_wide == ""] <- NA

# write.csv.gz(auc_unmatched_wide, file = "./data_munging/pozdeyev2016/merged_auc_vCHANGEDATE.3.csv.gz", row.names = FALSE, na = "NA")
```

```{r}
# test <- auc_unmatched_wide_test$InCCLE
# rm_numerics <- function(x) {
#   col_new <- c()
#   for(i in 1:length(x))
#     print(i)
#     x_unlist <- unlist(x[i])
#     print(x_unlist)
#     x_new <- ifelse(is.na(x_unlist[1]), 1000000, x_unlist[1])
#     print(x_new)
#     col_new <- c(col_new, x_new)
#     print(col_new)
#   return(col_new)
# }
# test2 <- rm_numerics(test)
```

@@

Column descriptions in merged_auc_v20190319.1.csv.gz:
- `Database`: the dataset from which the data comes from (either CCLE, CTRP, or GDSC)
- `Cell_line`: the name of the cancer cell line (CCL), harmonized across datasets by Pozdeyev et al. (2016)
- `Histology`: the general histology/tissue type of the CCL according to the CCLE annotations (see DepMap-2019q1-celllines_v2.csv file from the DepMap Data Portal)
- `sample_ID`: custom identifiers for this dataset that are concatenations of the `Database`, `Cell_line`, and `Histology`, separated by underscores
- `CCLE_ccl_name`: the cell line name using CCLE nomenclature, which is a concatenation of the cell line name with the CCLE-annotated histology in all caps, separated by underscores
- `Broad_ID`: the custom identifiers (formatted as ACH-######) used by the Broad's Cancer Dependency Map project to ease cell line name harmonization efforts 
- All other columns: compound and drug names that were screened in at least one CCL in at least one database
    - NA values indictate that compound wasn't screened in the respective CCL in the respective source dataset
    - Values are in AUC_IC50, the area under rhe dose response curve calculated from IC50 model

@@

>>>>

## CRISPR

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
```

@@

```{r, eval=FALSE}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
ccle_poz <- merge(ccl_histologies, ccle_poz, by = "Cell_line", all.y = TRUE)
ccle_poz$InCCLE <- "1"
ccle_poz$InCTRP <- ""
ccle_poz$InGDSC <- ""

@@

auc_unmatched <- bind_rows(ccle_poz, ctrp_poz, gdsc_poz)
auc_unmatched$WasDrugScreened <- 1
# auc_unmatched$AUC_EC50 <- as.numeric(as.character(auc_unmatched$AUC_EC50))
auc_unmatched$AUC_IC50 <- as.numeric(as.character(auc_unmatched$AUC_IC50))
# auc_unmatched$sample_ID <- with(auc_unmatched, paste(Database, Cell_line, Histology, sep = "_"))
auc_unmatched$Drug_name_Database <- with(auc_unmatched, paste(Drug_name, Database, sep = "_"))

@@

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_unmatched_wide <- dcast(auc_unmatched, Cell_line + Histology + CCLE_ccl_name + Broad_ID + InCCLE + InCTRP + InGDSC + WasDrugScreened ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_unmatched_wide[auc_unmatched_wide == 1000000] <- ""
auc_unmatched_wide <- auc_unmatched_wide %>% group_by(Cell_line, Histology, CCLE_ccl_name, Broad_ID, WasDrugScreened) %>% summarise_all(funs(trimws(paste(., collapse = ""))))
auc_unmatched_wide$InCCLE[auc_unmatched_wide$InCCLE == ""] <- 0
auc_unmatched_wide$InCTRP[auc_unmatched_wide$InCTRP == ""] <- 0
auc_unmatched_wide$InGDSC[auc_unmatched_wide$InGDSC == ""] <- 0
auc_unmatched_wide[auc_unmatched_wide == ""] <- NA
auc_unmatched_wide <- add_column(auc_unmatched_wide, WasCRISPRScreened = NA, .after = "WasDrugScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasCopyNumberData = NA, .after = "WasCRISPRScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasGeneExpressionData = NA, .after = "WasCRISPRScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasMutationData = NA, .after = "WasCRISPRScreened")

ccls_not_drugscreened <- filter(ccl_histologies, Cell_line %in% setdiff(ccl_histologies$Cell_line, auc_unmatched_wide$Cell_line))
ccls_not_drugscreened$WasDrugScreened <- 0

maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")

auc_all_ccls <- bind_rows(auc_unmatched_wide, ccls_not_drugscreened)
auc_all_ccls$HasMutationData <- with(auc_all_ccls, ifelse(Broad_ID %in% maf_df$Broad_ID, 1, 0))
auc_all_ccls$HasCopyNumberData <- with(auc_all_ccls, ifelse(Broad_ID %in% cn_melt$Broad_ID, 1, 0))
auc_all_ccls$HasGeneExpressionData <- with(auc_all_ccls, ifelse(Broad_ID %in% ge_melt$Broad_ID, 1, 0))
auc_all_ccls$WasCRISPRScreened <- with(auc_all_ccls, ifelse(Broad_ID %in% crispr_meta$Broad_ID, 1, 0))
auc_all_ccls$InCCLE <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCCLE))
auc_all_ccls$InCTRP <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCTRP))
auc_all_ccls$InGDSC <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InGDSC))
auc_all_ccls[, -c(1:4)] <- apply(auc_all_ccls[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all_ccls <- subset(auc_all_ccls, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all_ccls, file = "./data_munging/pozdeyev2016/merged_auc_v20190327.3.csv.gz", row.names = FALSE, na = "NA")
```

```{r}
# test <- as.data.frame(table(auc_all_ccls$Cell_line))
# test <- filter(auc_all_ccls, Cell_line %in% c("KMH2", "MS1", "TT"))[, 1:12]

length(unique(sapply(strsplit(as.character(colnames(auc_all_ccls)[13:length(colnames(auc_all_ccls))]), split = "_"), function(x) x[1])))
```

### README

@@

Column descriptions in merged_auc_v20190319.1.csv.gz:
- Cell_line: the name of the cancer cell line (CCL), harmonized across datasets by Pozdeyev et al. (2016)
- Histology: the general histology/tissue type of the CCL according to the CCLE annotations (see DepMap-2019q1-celllines_v2.csv file from the DepMap Data Portal)
- CCLE_ccl_name: the cell line name using CCLE nomenclature, which is a concatenation of the cell line name with the CCLE-annotated histology in all caps, separated by underscores
- Broad_ID: the custom identifiers (formatted as ACH-######) used by the Broad's Cancer Dependency Map project to ease cell line name harmonization efforts 
- WasDrugScreened: 1 means the cell was screened in at least one pharmacogenomic screens, and 0 means it was not screened in any of the pharmacogenomic screens
- WasCRISPRScreened: 1 means the cell line was screened in the Avana CRISPR screen, and 0 means it was not
- HasMutationData: 1 means the cell line had at least one recorded mutation in the MAF file (see file version below), and 0 means it was not annotated in the MAF file
- HasGeneExpressionData: 1 means gene expression data is available for the cell line (see file version below), and 0 means it does not have gene expression information
- HasCopyNumberData: 1 means copy number data is available for the cell line (see file version below), and 0 means it does not have copy number information
- InCCLE: 1 means the cell line was screened in CCLE, 0 means it was not
- InCTRP: 1 means the cell line was screened in CTRP, 0 means it was not
- InGDSC: 1 means the cell line was screened in GDSC, 0 means it was not
- All other columns: compound and drug names that were screened in at least one CCL in at least one database, formatted as Drug_DatabaseName
    - NA values indictate that compound wasn't screened in the respective CCL in the respective source dataset
    - Values are in AUC_IC50, the area under rhe dose response curve calculated from IC50 model

merged_auc_v20190319.1.csv.gz
merged_auc_v20190319.2.csv.gz:
- NAs in Histology column (excluding most of the cell lines indicated as being from a "collaborator" in CTRP) hand-curated by searches in DepMap, COSMIC, Cellosaurus, ATCC, canSAR Black, and Cell Model Passports
merged_auc_v20190327.3.csv.gz:
- Reformatted to have only unique cell lines in rows
- Cancer cell lines screened in more than one database have multiple AUC columns (column names are formatted as Drug_DatabaseName)
- Added information on binary presence/absencee of omic annotations for each cell line (i.e. mutations, copy number, gene expression)
- Added binary columns indicating presence in CCLE, CTRP, GDSC, drug screens in general (in 1, 2, or 3), and the Avana CRISPR screen
- Added cell lines that have at least one type of omic data but are not in the CRISPR or pharmacogenomic screens

@@

# Denise's GCN dataframes (Deprected)

## AUC

## Summary figures
